Title

### 2024-06-11 paper/main.tex
Updates intro

\begin{abstract}
	In Pearl's framework, the initial step for performing a causal analysis
	involves determining the causal structure between the variables from
	data typically in the form of a Directed Acyclic Graph (DAG) or an
	Structural Equation Model (SEM). This procedure, known as \emph{causal
	discovery}, has been extensively studied for both DAGs and SEMs. While
	numerous automated algorithms exist for estimating DAGs from datasets,
	their application in applied fields remain limited, possibly due to a
	lack of confidence in these algorithms stemming from them making
	obvious mistakes, and the difficulties in choosing the right algorithm
	for a given dataset. Consequently, in applied disciplines, these DAGs
	are predominantly constructed manually based on domain expertise. Due
	to being constructed by hand, it is vital that the fit of these models
	are tested against data and the models are potentially modified to
	acheive a better fit. On the other hand, in the field of SEMs models
	are typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} combined with expert knowledge can be used
	to guide this modification process. In the case of DAGs, although there
	are tests to assess the global fit of a model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying DAGs outputted by an algorithm. This
	modification process is based on using a (conditional) measure of
	association between variables in the model to assess potential edges
	that best explains unexplained correlation between variables which the
	researcher specifies the direction of this edge. In the case of
	continuous or discrete variables, the effect size measures of local
	tests can be used as this measure of association, however for mixed
	data there is no such measure. We present a measure of association
	based on canonical correlations for mixed data. We also present a
	graphical web-tool that can assist researchers in iteratively modifying
	and constructing their DAGs. We theoretically show that in the presence
	of an oracle, this iterative modification process is consistent in
	recovering the true DAG. Emprically, we show that using the mixed data
	measure of association, the modification process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

@@

A lot of questions in science center around a causal understanding between 
variables. \emph{Causal Discovery} methods aim to find these casual structure
among variables from a given dataset.

To perform any causal effect identification or estimation using Directed
Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
first construct a causal structure commonly known as DAG or path diagram. This
process of constructing the casual structure from data is known as \emph{Causal
Discovery}. In the field of DAGs, plenty of algorithms have been developed to
automatically construct DAGs from data. These algorithm takes different
approaches to causal discovery such as Constraint-based algorithms such as PC
and Fast Causal Infenrece (FCI) where the algorithm tries to match the
conditional independences in the data to the ones impled by the model,
score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
where the algorithm tries to find a model that has the best score given some
scoring metric, and continuous optimization where the problem is formulated as
a constrained optimization problem such as No Tears.

Even though we have many algorithms for casual discovery, the adaption of these
algorithms in applied fields remain limited. In these applied fields, researchers 
typically prefer to construct these DAGs manually based on their domain knowledge.
We hypothesize that this preference towards manual DAG construction stems from the 
following difficulties in applying these methods to real dataset.

\begin{enumerate}
	\item Algorithms output Markov Equivalence Class (MEC): Using observational
		data, any causal discovery algorithm can only output the MEC.
		MEC do not have all the edges oriented and to do further
		hypothesis testing or causal effect estimation, researchers
		typically have to rely on domain expertise to orient the
		remaining edges in the MEC.
	\item Difficulties in choosing an algorithm: To use any of these algorithms on 
		any practical dataset, users need to make a lot of choices. For
		example, for using constraint-based method, users need to
		choose a suitable condtional independence (CI) test, and
		hyperparameters for these tests. Similarly for score-based
		methods users need to choose the scoring metrics and their
		hyperparameters. Moreover, for practical implementation with
		reasonable runtime, users also need to choose the algorithm
		hyperparameters such as maximum number of conditioning
		variables to consider, max indegree of variables in the model. With
		no easy guides or method to evaluate these choices, it can be difficult.
	\item Lack of trust: Most of these automated algorithms can make very obvious 
		mistakes. Typically users have to test the outputs of the models and 
		manually make corrections to the output to get a usable model.
		This can lead to researchers not trusting these algorithms.
\end{enumerate}

On the other hand, in the field of Structural Equation Models (SEMs), models
are also typically constructed by hand. And a lot of focus has been put on
developing tools to help researchers in constructing these models. Researchers
typically come up with an initial model, which then they use the data and
expert knowledge to make modifications to. This process is commonly known as
\emph{Specification Search}. These tools can then be used to guide the
researchers in suggesting best ways to modify the model combined with the
expert knowledge. For example, \emph{modification indices} take the initial
drawn model, and make all possible modifications to the model to check which
modification improves the model fit the best. Based on this information, the
researcher can make modifications to the model and reiterate this process till
they get something that they are satisfied with.

>>>>

\begin{abstract}
	% TODO: Needs to be shortened.
	% In Pearl's framework, the initial step for performing a causal analysis
	% involves determining the causal structure between the variables from
	% data typically in the form of a Directed Acyclic Graph (DAG) or an
	% Structural Equation Model (SEM). This procedure, known as \emph{causal
	% discovery}, has been extensively studied for both DAGs and SEMs. While
	% numerous automated algorithms exist for estimating DAGs from datasets,
	% their application in applied fields remain limited, possibly due to a
	% lack of confidence in these algorithms stemming from them making
	% obvious mistakes, and the difficulties in choosing the right algorithm
	% for a given dataset. Consequently, in applied disciplines, these DAGs
	% are predominantly constructed manually based on domain expertise. Due
	% to being constructed by hand, it is vital that the fit of these models
	% are tested against data and the models are potentially modified to
	% acheive a better fit. On the other hand, in the field of SEMs models
	% are typically constructed manually through an iterative process of
	% assessing fit and modifying the model according to that. This process
	% is commonly known as \emph{Specification Search} where methods like
	% \emph{modification indices} combined with expert knowledge can be used
	% to guide this modification process. In the case of DAGs, although there
	% are tests to assess the global fit of a model by combining the tests
	% using implied conditional independencies of the model, but there are no
	% methods that can guide our modification process. In this paper, we
	% present a simple modification process aimed at helping researchers in
	% manually constructing or modifying DAGs outputted by an algorithm. This
	% modification process is based on using a (conditional) measure of
	% association between variables in the model to assess potential edges
	% that best explains unexplained correlation between variables which the
	% researcher specifies the direction of this edge. In the case of
	% continuous or discrete variables, the effect size measures of local
	% tests can be used as this measure of association, however for mixed
	% data there is no such measure. We present a measure of association
	% based on canonical correlations for mixed data. We also present a
	% graphical web-tool that can assist researchers in iteratively modifying
	% and constructing their DAGs. We theoretically show that in the presence
	% of an oracle, this iterative modification process is consistent in
	% recovering the true DAG. Emprically, we show that using the mixed data
	% measure of association, the modification process performs comparably to
	% PC and Hill-Climb Search algorithms if the user is able to correctly
	% identify the direction of the edge in one out of three cases.
\end{abstract}

@@

A lot of questions in life sciences center around a causal understanding
between variables. \emph{Causal Discovery} aims to find these casual structure
among random variables from a given dataset. The problem of causal discovery
has been studied both the in Directed Acyclic Graph (DAGs) and Structure
Equation Models (SEM) literatures. In the field of DAGs, many automated
algorithms have been developed to learn the casual structure form a given 
dataset such as constraint-based methods, score-based methods, and continuous
optimization based methods.

% To perform any causal effect identification or estimation using Directed
% Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
% first construct a causal structure commonly known as DAG or path diagram. This
% process of constructing the casual structure from data is known as \emph{Causal
% Discovery}. In the field of DAGs, plenty of algorithms have been developed to
% automatically construct DAGs from data. These algorithm takes different
% approaches to causal discovery such as Constraint-based algorithms such as PC
% and Fast Causal Infenrece (FCI) where the algorithm tries to match the
% conditional independences in the data to the ones impled by the model,
% score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
% where the algorithm tries to find a model that has the best score given some
% scoring metric, and continuous optimization where the problem is formulated as
% a constrained optimization problem such as No Tears.

However, the adaption of these algorithms in applied fields remain extremely
limited. In these applied fields, even though these algorithms exist,
researchers typically prefer to construct these DAGs manually based on their
domain knowledge. We hypothesize that this preference towards manual DAG
construction stems from the following difficulties in applying causal discovery
methods:

\begin{enumerate}
	\item Difficult to trust: Although asypotically these methods are 
		consistent, in the finite sample scenario they can end up
		making obvious mistakes. As a result of this, researchers
		have to rely on posthoc modification of the output of these
		algorithms using either their domain knowledge or assisted
		by testing methods.
	\item Algorithms output the Markov Equivalence Class (MEC): With any
		causal discovery algorithm, we can only get to the MEC. However,
		most of the methods developed for downstream tasks such 
		as identification or casual effect estimation assume the
		knowledge of a DAG. To orient these MECs to a DAG, researchers
		have to either rely on their domain expertise, or make further
		assumption to utilize a pairwise orientation method.
	\item Difficulties in choosing an algorithm: All the algorithms
		have the same goal with all of them having similar asympottic
		properties. However, the finite sample properties of these
		methods vary depending on the choices that we make. For
		example, the choice of the conditional independence (CI) test
		being used for constraint-based method depends on the dataset,
		similarly for the scoring method for score-based methods. With
		no obvious way to test the impact/performance of these choices
		on a given dataset, it becomes quite hard to chose the optimal
		algorithm for a given dataset.
\end{enumerate}

On the other hand, in the field of Structural Equation Models (SEMs), a lot
more focus has been put on methods to help researchers draw models manually
along with integrating their expertise. Researchers typically come up with an
initial model based on their domain knowledge, and then use various tools to
guide them in making modifications to this initial model. This process of
constructing models is commonly known as \emph{Specification Search}. A
commonly used method for suggesting potential modifications to the SEM is
\emph{modification indices}. Given a specified model, modification indices
tries to find potential modifications to the model that improve the model's fit
to the data best. Based on the suggestions from this method along with their
domain expertise, the researcher can then choose the most optimal modification
for their model.
